---
title: "Data overview plots"
author: "Mélanie Jean"
date: "5/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/My documents/CitizenScienceOnLockdown/Data")
```
Open useful libraries. 
```{r}
library(tidyverse)
```

Import data
```{r}
iNat=read.table("../Data//04_Inat_from_clipping_NoDuplicates-csv.tsv",quote= "\"", fill=T, sep='\t',comment.char = "", na.strings=c("NA", ""), header=T)
```
Some observations from the variable 'city' for Ottawa are in Ontario and some are in Quebec. Same for Gatineau. The issue is resolved when using the variable Ggrphc_n instead. 

Plot observations by city x year to see whether the cities we selected actually use iNat. 
```{r}
CityTotal=iNat%>%
# mutate(Ggrphc_n = plyr::revalue(Ggrphc_n, c("Montreal"="Montreal")))%>%
 group_by(Prvnc__, Ggrphc_n, year)%>%
  count()

CityTotal$Province=as.factor(CityTotal$Prvnc__)

CityTotal$city=as.factor(CityTotal$Ggrphc_n)
CityTotal$city <- factor(CityTotal$Ggrphc_n, levels = c("Vancouver","Surrey", "Calgary","Edmonton", "Saskatoon","Winnipeg","London","Hamilton", "Brampton", "Mississauga","Toronto", "Markham","Vaughan", "Ottawa", "Gatineau" , "Laval" , "Montreal","Longueuil",  "Quebec",  "Halifax"))
```
Plot with different colors for different provinces. Quebec-blue, alberta-yellow, bc-green. 
-> Weird some Ottawa and Gatineau observations are in the wrong province...
```{r}
p=ggplot(data=CityTotal, aes(x=year, y=n))+
	geom_bar(stat="identity", aes(fill=as.factor(Province)))+
	facet_wrap(~as.factor(city))+
	scale_y_continuous(name="Observation counts")+
	scale_fill_manual(name="Province", values=c("#F8766D","#FB61D7","#53B400","#A58AFF", "#00C094" ,"#00B6EB",  "#C49A00"))+
    theme_bw()
p

#00C094 teal
#FB61D7 pink
#F8766D coral
#53B400 green
#C49A00 yellow-ish
#A58AFF purple
#00B6EB blue
```

Import population data
```{r}
Pop20=read.csv("01_Population_top20.csv", header=T)

```
Add populaiton total and density to the city total table
```{r}
Pop20.cols=Pop20%>%
  select(Geographic.name, Population..2016, Population.density.per.square.kilometre..2016)%>%
  rename(city=Geographic.name) #Change the column names for something more manageable

CityTotal2=left_join(CityTotal, Pop20.cols)%>%
  mutate(Observations_capita=n/Population..2016)

CityTotal2$city <- factor(CityTotal2$city, levels = c("Vancouver","Surrey", "Calgary","Edmonton", "Saskatoon","Winnipeg","London","Hamilton", "Brampton", "Mississauga","Toronto", "Markham","Vaughan", "Ottawa", "Gatineau" , "Laval" , "Montréal","Longueuil",  "Quebec",  "Halifax"))

p2=ggplot(data=CityTotal2, aes(x=year, y=Observations_capita))+
  geom_bar(stat="identity", aes(fill=as.factor(city)))+
  facet_wrap(~as.factor(city))+
  scale_y_continuous(name="Observation per capita")+
  theme_bw()+
 theme(legend.position = "none")
p2
```

